Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2019-12-11 paper/manuscript.tex
finishing work through manuscript

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1_what_is_alchemy/Figure.pdf}   
    \caption{\textbf{Illustration of the types of free energies differences that can be calculated using alchemical free energy methods.} A: Change in free energy due to a conformational changes of the moelcule. B: Partition coefficient such as $\log P$ or $\log D$ can be computed using alchemical free energy methods. Here, as an example the partition coefficient between methanol and water is shown. C: Free energy difference associated with the insertion of a molecule into a membrane. D: Effect of mutations of protein or host residues on free energies of binding.  E: Absolute free energies of binding of a small molecule to a host (e.g. protein), F: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.
    \label{fig:fig1_what_is_alchemy}
    }
\end{figure}

@@

\begin{equation}
    \Delta G_{\mathrm{bind}} = -RT\ln\frac{\int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}}{\int_\mathrm{unbound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}}
\end{equation}
Or more generally in a simplified form the free energy difference between the bound and unbound states is:
\begin{equation}
\Delta f \equiv f(\mathrm{bound}) - f(\mathrm{unbound}) = -\ln\frac{Z(\mathrm{unbound})}{Z({\mathrm{bound}})},
\end{equation}
where $Z = \int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}$ represents the configurational integral. 
How can we compute $\Delta G_{\mathrm{bind}}$ from a molecular dynamics simulation? Running a direct simulation of the binding events is too computationally costly, meaning we cannot easily estimate the configurational integrals of the bound and unbound state from the simulation. This is where alchemical free simulations come in. We can write down a thermodynamic cycle as seen in ~\ref{fig:fig3_binding-thermodynamic-cycle}, for a relative free energy simulation. Based on this cycle we can compute the relative free energy of binding between two compounds $A$ and $B$, by going along the arrow of $\Delta G_{\mathrm{bound}}$, where we introduce a second compound that is similar to the first one. Now $\Delta G_{\mathrm{bound}}$ can be expressed as a ratio of partition functions of toluene (A) and benzyl alcohol (B). 
The defining characteristic of alchemical free energy calculations is the use of a series of alchemically-modified potential functions $U(x; \lambda)$ in which an alchemical parameter $\lambda$ modulates interactions in a manner that cannot occur in real chemical systems.
One or more simulations are used to collect data from a multitude of alchemical states to compute a free energy difference between a chemical state ($\lambda_0$) and another chemical or alchemical reference state ($\lambda_1$),
\begin{eqnarray}
\Delta f &\equiv& f(\lambda_1) - f(\lambda_0) = - \ln \frac{Z(\lambda_1)}{Z(\lambda_0)} , \label{equation:dimensionless-free-energy-difference}
\end{eqnarray}
where the dimensionless free energy $f(\lambda) \equiv \beta F(\lambda)$ is given in terms of partition functions $Z(\lambda)$,
\begin{eqnarray}
Z(\lambda) &=& \int dx \, e^{-u(x; \lambda)} .
\label{equation:partition-function-definition}
\end{eqnarray}
Here, the inverse thermal energy $\beta \equiv (k_B T)^{-1}$ where $k_B$ is the Boltzmann constant and $T$ is the absolute temperature, and the \emph{reduced potential} $u(x; \lambda)$~\cite{shirts2008statistically} is generally given by a trace over thermodynamic parameters with their conjugate dynamical variables,
\begin{eqnarray}
u(x;\lambda) &\equiv& \beta \left[ U(x;\lambda) + p \, V(x) + \sum_{i=1}^N \mu_i \, N_i(x) + \cdots \right] . \label{equation:reduced-potential}
\end{eqnarray}
Here, the collection of thermodynamic and alchemical parameters $\theta \equiv \{\beta, \lambda, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
The physical transformation of interest---such as a protein-ligand association process, or transfer free energy among phases---may involve a thermodynamic cycle that requires several free energy differences $\Delta f$ to be computed in order to produce an estimate of the overall desired quantity.

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig2_therm_cyc/Figure.pdf}
    \caption{{\bf Thermodynamic cycle for computing the relative free energy of binding ($\Delta \Delta G$) between two related small molecules to a supramolecular host.}
    The relative binding free energy difference between two small molecules, $\Delta \Delta G_{\mathrm{bind}, A \rightarrow B} \equiv \Delta G_{\mathrm{bind}, B} - \Delta G_{\mathrm{bind}, A}$---here benzyl alcohol (top) to toluene (bottom)---can be computed as a difference between two alchemical transformations, $\Delta G_\mathrm{bound} - \Delta G_\mathrm{solvated}$, where $\Delta G_\mathrm{bound}$ represents the free energy change of transforming $A \rightarrow B$ in complex and $\Delta G_\mathrm{solvated}$ the free energy change of transforming $A \rightarrow B$ in solvent.}
    \label{fig:fig_binding-thermodynamic-cycle}
\end{figure}

@@

\todo[inline, color={green!20}]{@everyone please review this section}

@@

Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. An overview of benchmark datasets and case studies can be found in section \ref{sec:benchmark} at the end of this document.
\todo[inline,color={green!20}]{JDC: We should probably provide a short list of key benchmark references here, broken down by force field, and keep this up to date. GT: would a table, at the end of the document be an easier way to do this, only requiring some brief comments in the text? ASJSM: I like the table idea. MK: Feel free to remove this todo item if you are happy with the table provided. I refrained from just listing lots of case studies, as I think we should highlight the recommended datasets which should be used for benchmarking. Nevertheless, I have cited an review article containing such a summary as well as a recent publication comparing FEP+ and Gromacs in the brief text below the table.}
However, the domain of applicability is a significant concern~\cite{sherborne2016collaborating, cournia2017relative}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

It is worth noting that the accuracy of free energy calculations is highly variable across different protein targets, and likely across different ligand chemotypes as well. For instance, FEP+ with OPLS3 achieves an RMSE of 0.62 kcal/mol for a set of 21 compounds binding to JNK1 kinase, yet only an RMSE of 1.05 kcal/mol for a set of 34 compounds binding to P38$\alpha$ kinase~\cite{harder2016opls3}. Furthermore, perturbations for the same chemotype in different pockets of the BACE enzyme gave varied errors~\cite{keranen2017acylguanidine}
\todo[inline,color={green!20}]{JDC: It's important to be clear on what error is being reported: $\Delta G$ after shifting by a constant to minimize the RMSE, unshifted $\Delta G$, $\Delta \Delta G$ of computed edges, or $\Delta \Delta G$ of all edges. Note that computed edge $\Delta \Delta G$ RMSE significantly underestimates the error of all edge RMSE: \url{https://github.com/jchodera/jacs-dataset-analysis} ASJSM: Can we incorporate your analysis in this section as an example?}
We thus recommend that retrospective study for a particular target and a particular chemical series be performed to establish the relevant accuracy of free energy calculations for each particular application case.
%

@@

\todo[inline]{Discuss implications of the SAMPL6 SAMPLing paper~\cite{rizzi2019sampl6} on reproducibility between simulation packages and potential sensitivity to other calculation details.}

@@

\todo[inline, color={blue!20}]{@DM: A plot of this might be useful. I believe you have used/generated one of these before.}

@@

\todo[inline, color={green!20}]{@everyone: Might be nice to do an example calculation X perturbation on GTX 1080s on a AWS cluster cost \$\$X. Synthesis costs \$\$Y. }

@@

\section{How should alchemical simulations be applied to drug discovery?}

\todo[inline]{JDC: Should this be a Section or Subsection?}
\todo[inline]{JDC: This section needs some structure (subsections, paragraph) to break up the "wall of text".}

@@

In summary, the successful use of alchemical calculations in drug discovery requires working in the domain of applicability, using a high quality X-ray structure of the target compounds, and testing the approach retrospectively to ensure the system setup is well-behaved. Always assess the confidence in the predictions and communicate this when discussing with experimentalists. Consider performing repeat calculations for at least some of the perturbations in the study. In a drug discovery setting it is inevitable to consider larger virtual libraries, in these cases again evaluate if shorter simulation time per $\lambda$ window can be sufficient for a ‘virtual screening’ mode of application. It is expected that future tools may allow efficient exploration of large virtual libraries. There are many accounts of success of alchemical FE calculations, the methods show good performance towards the goal of binding energy prediction. However, it is important to have realistic expectations. Structure based drug design projects are often capable of improving potency relatively quickly and the range of activity narrows to just two-to-three log units. It may seem hard to impact with substantially different, more potent, stand-out compounds in this scenario, but binding energy prediction can still be extremely useful for ensuring activity is maintained as other properties are optimized. An interesting cost benefit analysis has been performed, see drug discovery strategy related discussions of Mobley Abel and co. From a drug discovery point of view, FEP is expanding its domain of applicability, there are reports of success using homology models and GPCRs for instance, as well as enabling charge change and scaffold hopping, but these systems are undoubtedly more difficult.  In the meantime the use cases are expanding to resistance prediction, disease mutation prediction, solubility prediction – an exciting future for alchemical FE calculations. 
%
%
%
%%%%%%%%%%%%
% Step 1   %
%%%%%%%%%%%%
%
\section{Simulation prerequisites}
\label{sec:step1}
\todo[inline,color={green!20}]{@everyone: please review}
The alchemical free energy protocols, that are discussed in Section~\ref{sec:step2}, are defined for a specific type of free energy calculation, i.e. a free energy of binding or a free energy of hydration. Depending on the type of simulations different considerations have to be made for ligands, solvent, and host molecules (in the case of the estimation of free energies of binding).
\subsection*{Free energies of binding}
In principle in the limit of sufficient conformational sampling, the free energy changes estimated from an alchemical free energy calculation should be independent of the system's initial coordinates. However, in practice because simulations of thermodynamic states are of finite duration (typically currently 1-100 ns per state) this is only true for certain classes of alchemical free energy calculations such as relative or absolute free energies of hydration of small and relatively rigid organic molecules. Host-guest or protein-ligand complexes typically exhibit slowly relaxing degrees of freedom that exceed significantly the duration of an alchemical free energy calculation. It is therefore generally important to carefully select input coordinates to obtain satisfactory results. 
The following questions may be relevant before diving into the simulation setup. 
\begin{itemize}
    \item Do I have one or multiple good host structures? (e.g. a good resolution X-ray crystal of the protein target)
    \item Should I include buried waters, or other small molecules that can be found in an X-ray structure.
    \item Are my ligands part of a congeneric series? (simple R group substitutions around the same scaffold)
    \item Do I have information on one or all of the ligands binding sites (e.g. a X-ray structure)
\end{itemize}

As for any simulation, care should be taken in selecting available X-ray structures in the protein databank and potentially choosing multiple structures initially, to account for variability in receptor conformations as well as accuracy of available X-ray structures. Typically, clustering of receptor structures can be used to identify different receptor conformations near the binding site, as well as assessing relevant side chain placements from the X-ray structure, see for example~\cite{mey2016blinded}. In terms of set up and other choices following general best practice guidelines is advisable~\cite{braun2018best}.

@@

Generally, binding modes within congeneric series are conserved~\cite{}, however, exceptions exist~\cite{}, and weaker binding compounds are more likely to adopt different binding modes (REFS). There may also be ambiguity about the relative orientation of substituents. A known issue involves a 180 degree flip in the dihedral angle of a 6-membered aromatic ring leading to a different spatial position of ortho- or meta- substituents that otherwise should overlap within a series. The 180 degree flip of the ring may not occur enough during simulations to overcome the starting configuration bias. Another scenario may be equatorial and axial substituted saturated rings (e.g. cyclohexane derivatives). This situation may be addressed by explicitly modelling different binding modes of the same ligand and combining later computed free energy differences for different binding modes into a relative free energies of binding~\cite{kaus2015how}.

Congeneric series can contain stereoisomers or enantiomers, making sure that you model the correct one is important. For racemates, the relative abundance of each stereoisomer is normally not known. Therefore, the experimental activity associated with just one conformation is more uncertain. However, the modeling typically uses just the bioactive conformation that best fits the active site. Clearly this introduces the chance for higher error versus experiment. Nevertheless, if all compounds in the congeneric series are racemic, originating from similar synthetic procedures with an expected similar abundance of stereoisomers, then the differences may cancel and the trend in calculated and observed binding energies may be robust. Nevertheless, we can see that care and further testing is needed in this scenario, and the quality of the predictions may suffer.

Another aspect to look out for are binding site water molecules that may form water mediated protein-ligand interactions, particularly in situations where exchange with bulk water may be slow on the simulation timescales. This happens typically in buried binding site. Overlaying multiple protein X-ray structures can identify conserved or just additional water molecules that can be useful to include. In cases where water molecules are known to play an important role in the binding, software implementations that use water sampling facilitated by Grand Canonical Monte Carlo methods may be useful. For examples see HSP90 as formed part of the D3R grand challenge 2015~\cite{mey2016blinded},A2A~\cite{brucemacdonald2018ligand}, MUP~\cite{ross2015water}, and others~\cite{michel2009energetics}.

How to align congeneric series:
\todo[inline, color=green!20]{Question: How do people align congeneric ligand series?}

Input coordinates for a congeneric series may be generated by docking calculations, or by ligand alignment using maximum common substructure algorithms. The latter tends to produce alignments that are more conserved and more consistent free energy changes across a dataset, but will struggle to yield reasonable results for relative binding free energy calculations that involve a significant binding mode rearrangement. This may also lead to steric clashes with the receptor coordinates of the reference ligand if structural rearrangements are needed to accommodate different members of the congeneric series. Small steric clashes may be resolved during subsequent simulation equilibration prior to data collection but there is a risk that the complex relaxes to an alternative metastable state. 
\todo[inline, color=green!20]{Just a note as a reminder: Chirality!}
An additional consideration arises for single topology relative free energy calculations. In this class of alchemical free energy calculations it is necessary to generate a molecular topology that may describe the initial and final states of the perturbation (see 7.1.1). In cases where the end-states have high topological similarity and high structural overlap this is relatively straightforward and typically handled by use of maximum common substructure calculations. In situations where the end-state topologies differ significantly, or there is relatively little spatial overlap between the two end-states some user intervention may be necessary to produce a satisfactory input topology.

@@

%
%
%
%%%%%%%%%%%%
% Step 1   %
%%%%%%%%%%%%
%
\section{What simulation protocol should I choose? -- Step 2}
\label{sec:step2}
Alchemical free energy calculations can be grouped into two main categories, ``absolute'' and ``relative'' \footnote{The distinction is a bit of a misnomer, since both compute ratios of partition functions relative to another state, and neither computes an absolute free energy.}, which differ in whether they compute properties for a single molecule (absolute) or compare properties of different, usually closely related, molecules (relative).
To use binding as a concrete example, in absolute binding free energy calculations, one computes the binding free energy of a ligand to an individual receptor relative to a standard reference concentration.
In contrast, in relative binding free energy calculations, one compares the binding free energy of two related inhibitors to determine the potency difference.
\subsection*{Absolute and relative free energy calculations have some differences}
Many of the issues of setup and protocol for alchemical calculations are common, but some are different between absolute and relative calculations, so before treating the common elements we consider the protocol differences.

@@

\paragraph{Topologies} A critical first step in relative calculations is to select an approach to these calculations, determining whether to use a \emph{dual topology}, \emph{single topology}, or \emph{hybrid topology} approach to relative calculations.
\begin{figure}
    \includegraphics[width=0.95\columnwidth]{figures/fig3_topol/Figure.pdf}
    \caption{top: single topology bottom: dual topology Figure adapted from \url{http://www.alchemistry.org/wiki/Constructing_a_Pathway_of_Intermediate_States}}
    \label{fig:fig_topology}
\end{figure} 
The distinction between these can be illustrated by considering a hypothetical transformation from molecule A to molecule B, where both atoms share a common substructure but differ in their substituents; e.g. consider a transformation of benzene (A) to benzyl alcohol (B)~\ref{fig:fig4_topology}.
In this case the common substructure is the benzene ring, though perhaps may be larger depending on how it is defined, as we discuss below.
In single topology calculations, the overall transformation is set up to involve as few additional atoms as possible, so benzene would be typically changed into benzyl alcohol by changing one of the hydrogens into a carbon. To this new carbon you also find non-interacting atoms called ``dummy atoms'' (retaining their bonded interactions but not interacting with the rest of the system), which will become the two hydrogens attached to the new carbon and the OH group. Bond parameters as well as partial charges are adjusted accordingly. 
Thus, in a single topology calculation, atoms may change their type so relatively few dummy atoms are created. This is illustrated in fig.~\ref{fig:fig4_topology} (a). 
In contrast, in a dual topology free energy calculation, no atoms are allowed to change type~\cite{shirts2012best}. This means that the benzene to benzyl alcohol transformation involves starting with benzene plus the non-interacting dummy atoms making up the hydroxy methyl group, then passing through an intermediate state where atoms which are becoming dummy atoms or ceasing being dummy atoms are partially interacting (this state may or may not be well defined~\cite{mobley2014blind}  and culminating in a state where benzyl alcohol is present along with the additional dummy atoms which was previously a corresponding hydrogen of the benzene. Fig.~\ref{fig:fig4_topology} (b) depicts how such a dual topology works.

>>>>

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1_what_is_alchemy/Figure.pdf}   
    \caption{\textbf{Illustration of the types of free energies differences that can be calculated using alchemical free energy methods.} A: Change in free energy due to a conformational changes of the moelcule. B: Partition coefficient such as $\log P$ or $\log D$ can be computed using alchemical free energy methods. Here, as an example the partition coefficient between methanol and water is shown. C: Free energy difference associated with the insertion of a molecule into a membrane. D: Effect of mutations of protein or host residues on free energies of binding.  E: Absolute free energies of binding of a small molecule to a host (e.g. protein), F: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.
    \label{fig:fig_what_is_alchemy}
    }
\end{figure}

@@

\begin{equation}
    \Delta G_{\mathrm{bind}} = -RT\ln\frac{\int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}}{\int_\mathrm{unbound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}}
\end{equation}
Or more generally in a simplified form the free energy difference between the bound and unbound states is:
\begin{equation}
\Delta f \equiv f(\mathrm{bound}) - f(\mathrm{unbound}) = -\ln\frac{Z(\mathrm{unbound})}{Z({\mathrm{bound}})},
\end{equation}
where $Z = \int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}$ represents the configurational integral. 
How can we compute $\Delta G_{\mathrm{bind}}$ from a molecular dynamics simulation? Running a direct simulation of the binding events is too computationally costly, meaning we cannot easily estimate the configurational integrals of the bound and unbound state from the simulation. This is where alchemical free simulations come in. We can write down a thermodynamic cycle as seen in ~\ref{fig:fig_binding_thermodynamic_cycle}, for a relative free energy simulation. Based on this cycle we can compute the relative free energy of binding between two compounds $A$ and $B$, by going along the arrow of $\Delta G_{\mathrm{bound}}$, where we introduce a second compound that is similar to the first one. Now $\Delta G_{\mathrm{bound}}$ can be expressed as a ratio of partition functions of toluene (A) and benzyl alcohol (B). 
The defining characteristic of alchemical free energy calculations is the use of a series of alchemically-modified potential functions $U(x; \lambda)$ in which an alchemical parameter $\lambda$ modulates interactions in a manner that cannot occur in real chemical systems.
One or more simulations are used to collect data from a multitude of alchemical states to compute a free energy difference between a chemical state ($\lambda_0$) and another chemical or alchemical reference state ($\lambda_1$),
\begin{eqnarray}
\Delta f &\equiv& f(\lambda_1) - f(\lambda_0) = - \ln \frac{Z(\lambda_1)}{Z(\lambda_0)} , \label{equation:dimensionless-free-energy-difference}
\end{eqnarray}
where the dimensionless free energy $f(\lambda) \equiv \beta F(\lambda)$ is given in terms of partition functions $Z(\lambda)$,
\begin{eqnarray}
Z(\lambda) &=& \int dx \, e^{-u(x; \lambda)} .
\label{equation:partition-function-definition}
\end{eqnarray}
Here, the inverse thermal energy $\beta \equiv (k_B T)^{-1}$ where $k_B$ is the Boltzmann constant and $T$ is the absolute temperature, and the \emph{reduced potential} $u(x; \lambda)$~\cite{shirts2008statistically} is generally given by a trace over thermodynamic parameters with their conjugate dynamical variables,
\begin{eqnarray}
u(x;\lambda) &\equiv& \beta \left[ U(x;\lambda) + p \, V(x) + \sum_{i=1}^N \mu_i \, N_i(x) + \cdots \right] . \label{equation:reduced-potential}
\end{eqnarray}
Here, the collection of thermodynamic and alchemical parameters $\theta \equiv \{\beta, \lambda, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
The physical transformation of interest---such as a protein-ligand association process, or transfer free energy among phases---may involve a thermodynamic cycle that requires several free energy differences $\Delta f$ to be computed in order to produce an estimate of the overall desired quantity.

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig2_therm_cyc/Figure.pdf}
    \caption{{\bf Thermodynamic cycle for computing the relative free energy of binding ($\Delta \Delta G$) between two related small molecules to a supramolecular host.}
    The relative binding free energy difference between two small molecules, $\Delta \Delta G_{\mathrm{bind}, A \rightarrow B} \equiv \Delta G_{\mathrm{bind}, B} - \Delta G_{\mathrm{bind}, A}$---here benzyl alcohol (top) to toluene (bottom)---can be computed as a difference between two alchemical transformations, $\Delta G_\mathrm{bound} - \Delta G_\mathrm{solvated}$, where $\Delta G_\mathrm{bound}$ represents the free energy change of transforming $A \rightarrow B$ in complex and $\Delta G_\mathrm{solvated}$ the free energy change of transforming $A \rightarrow B$ in solvent.}
    \label{fig:fig_binding_thermodynamic_cycle}
\end{figure}

@@

\todo[inline, color={green!20}]{ASJSM: @everyone please review this section}

@@

Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. A small selection of benchmark datasets and case studies can be found in section~\ref{sec:benchmark} at the end of this document.
However, the domain of applicability is a significant concern~\cite{sherborne2016collaborating, cournia2017relative}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

It is worth noting that the accuracy of free energy calculations is highly variable across different protein targets, and likely across different ligand chemotypes as well. For instance, FEP+ with OPLS3 achieves an RMSE of 0.62 kcal/mol for a set of 21 compounds binding to JNK1 kinase, yet only an RMSE of 1.05 kcal/mol for a set of 34 compounds binding to P38$\alpha$ kinase~\cite{harder2016opls3}. Furthermore, perturbations for the same chemotype in different pockets of the BACE enzyme gave varied errors~\cite{keranen2017acylguanidine}
\todo[inline,color={green!20}]{JDC: It's important to be clear on what error is being reported: $\Delta G$ after shifting by a constant to minimize the RMSE, unshifted $\Delta G$, $\Delta \Delta G$ of computed edges, or $\Delta \Delta G$ of all edges. Note that computed edge $\Delta \Delta G$ RMSE significantly underestimates the error of all edge RMSE: \url{https://github.com/jchodera/jacs-dataset-analysis} ASJSM: Can we incorporate your analysis in this section as an example? Maybe this is something that needs to be addressed properly in the analysis section and only referenced here?}
We thus recommend that retrospective study for a particular target and a particular chemical series be performed to establish the relevant accuracy of free energy calculations for each particular application case.
%

@@

\todo[inline,color={green!20}]{ASJSM: Discuss implications of the SAMPL6 SAMPLing paper~\cite{rizzi2019sampl6} on reproducibility between simulation packages and potential sensitivity to other calculation details. @jmichel, @dmobley}

@@

@@

\todo[inline, color={green!20}]{ASJSM: @everyone: Might be nice to do an example calculation X perturbation on GTX 1080s on a AWS cluster cost \$\$X. Synthesis costs \$\$Y. }

@@

\section{How should alchemical simulations be applied to drug discovery?}
\todo[inline, color={green!20}]{ASJSM: Merge/tidy of what can I expect from my simulation, drug discovery aspects and simulation prerequisites, there are some duplications there at the moment, happy for restructuring ideas.  }

@@

In summary, the successful use of alchemical calculations in drug discovery requires working in the domain of applicability, using a high quality X-ray structure of the target compounds, and testing the approach retrospectively to ensure the system setup is well-behaved. Always assess the confidence in the predictions and communicate this when discussing with experimentalists. Consider performing repeat calculations for at least some of the perturbations in the study. In a drug discovery setting it is inevitable to consider larger virtual libraries, in these cases again evaluate if shorter simulation time per $\lambda$ window can be sufficient for a ‘virtual screening’ mode of application. It is expected that future tools may allow efficient exploration of large virtual libraries. There are many accounts of success of alchemical FE calculations, the methods show good performance towards the goal of binding energy prediction. However, it is important to have realistic expectations. Structure based drug design projects are often capable of improving potency relatively quickly and the range of activity narrows to just two-to-three log units. It may seem hard to impact with substantially different, more potent, stand-out compounds in this scenario, but binding energy prediction can still be extremely useful for ensuring activity is maintained as other properties are optimized. An interesting cost benefit analysis has been performed, see drug discovery strategy related discussions of Mobley Abel and co. From a drug discovery point of view, FEP is expanding its domain of applicability, there are reports of success using homology models and GPCRs for instance, as well as enabling charge change and scaffold hopping, but these systems are undoubtedly more difficult.  In the meantime the use cases are expanding to resistance prediction, disease mutation prediction, solubility prediction – an exciting future for alchemical FE calculations. 
%
%
%
%%%%%%%%%%%%
% Step 1   %
%%%%%%%%%%%%
%
\section{Simulation prerequisites}
\label{sec:step1}

The alchemical free energy protocols, that are discussed in Section~\ref{sec:step2}, are defined for a specific type of free energy calculation, i.e. a free energy of binding or a free energy of hydration. Depending on the type of simulations different considerations have to be made for ligands, solvent, and host molecules (in the case of the estimation of free energies of binding).
\subsection*{Free energies of binding}
In principle in the limit of sufficient conformational sampling, the free energy changes estimated from an alchemical free energy calculation should be independent of the system's initial coordinates. However, in practice because simulations of thermodynamic states are of finite duration (typically currently 1-100 ns per state) this is only true for certain classes of alchemical free energy calculations such as relative or absolute free energies of hydration of small and relatively rigid organic molecules. Host-guest or protein-ligand complexes typically exhibit slowly relaxing degrees of freedom that exceed significantly the duration of an alchemical free energy calculation. It is therefore generally important to carefully select input coordinates to obtain satisfactory results. 
The following questions may be relevant before diving into the simulation setup. 
\begin{itemize}
    \item Do I have one or multiple good host structures? (e.g. a good resolution X-ray crystal of the protein target)
    \item Should I include buried waters, or other small molecules that can be found in an X-ray structure.
    \item Are my ligands part of a congeneric series? (simple R group substitutions around the same scaffold)
    \item Do I have information on one or all of the ligands binding sites (e.g. a X-ray structure)
\end{itemize}

\paragraph{Are there good X-ray structures available?}
As for any simulation, care should be taken in selecting available X-ray structures in the protein databank and potentially choosing multiple structures initially, to account for variability in receptor conformations as well as accuracy of available X-ray structures. Typically, clustering of receptor structures can be used to identify different receptor conformations near the binding site, as well as assessing relevant side chain placements from the X-ray structure, see for example~\cite{mey2016blinded}. In terms of set up and other choices following general best practice guidelines is advisable~\cite{braun2018best}.

@@

\paragraph{How to deal with different binding modes?}
Generally, binding modes within congeneric series are conserved~\cite{}, however, exceptions exist~\cite{}, and weaker binding compounds are more likely to adopt different binding modes (REFS). There may also be ambiguity about the relative orientation of substituents. A known issue involves a 180 degree flip in the dihedral angle of a 6-membered aromatic ring leading to a different spatial position of ortho- or meta- substituents that otherwise should overlap within a series. The 180 degree flip of the ring may not occur enough during simulations to overcome the starting configuration bias. Another scenario may be equatorial and axial substituted saturated rings (e.g. cyclohexane derivatives). This situation may be addressed by explicitly modelling different binding modes of the same ligand and combining later computed free energy differences for different binding modes into a relative free energies of binding~\cite{kaus2015how}.

\paragraph{Have you considered sterioisomers and enationmers?}
Congeneric series can contain stereoisomers or enantiomers, making sure that you model the correct one is important. For racemates, the relative abundance of each stereoisomer is normally not known. Therefore, the experimental activity associated with just one conformation is more uncertain. However, the modeling typically uses just the bioactive conformation that best fits the active site. Clearly this introduces the chance for higher error versus experiment. Nevertheless, if all compounds in the congeneric series are racemic, originating from similar synthetic procedures with an expected similar abundance of stereoisomers, then the differences may cancel and the trend in calculated and observed binding energies may be robust. Nevertheless, we can see that care and further testing is needed in this scenario, and the quality of the predictions may suffer.

\paragraph{Conserved binding site waters can play an important role in binding free energies}
Another aspect to look out for are binding site water molecules that may form water mediated protein-ligand interactions, particularly in situations where exchange with bulk water may be slow on the simulation timescales. This happens typically in buried binding site. Overlaying multiple protein X-ray structures can identify conserved or just additional water molecules that can be useful to include. In cases where water molecules are known to play an important role in the binding, software implementations that use water sampling facilitated by Grand Canonical Monte Carlo methods may be useful. For examples see HSP90 as formed part of the D3R grand challenge 2015~\cite{mey2016blinded},A2A~\cite{brucemacdonald2018ligand}, MUP~\cite{ross2015water}, and others~\cite{michel2009energetics}.

\paragraph{How to align congeneric series}
Input coordinates for a congeneric series may be generated by docking calculations, or by ligand alignment using maximum common substructure algorithms. The latter tends to produce alignments that are more conserved and more consistent free energy changes across a dataset, but will struggle to yield reasonable results for relative binding free energy calculations that involve a significant binding mode rearrangement. This may also lead to steric clashes with the receptor coordinates of the reference ligand if structural rearrangements are needed to accommodate different members of the congeneric series. Small steric clashes may be resolved during subsequent simulation equilibration prior to data collection but there is a risk that the complex relaxes to an alternative metastable state.

An additional consideration arises for single topology relative free energy calculations. In this class of alchemical free energy calculations it is necessary to generate a molecular topology that may describe the initial and final states of the perturbation (see~\ref{}). In cases where the end-states have high topological similarity and high structural overlap this is relatively straightforward and typically handled by use of maximum common substructure calculations. In situations where the end-state topologies differ significantly, or there is relatively little spatial overlap between the two end-states some user intervention may be necessary to produce a satisfactory input topology.

@@

%
%
%
%%%%%%%%%%%%
% Step 1   %
%%%%%%%%%%%%
%
\section{What simulation protocol should I choose? -- Step 2}
\label{sec:step2}
Alchemical free energy calculations can be grouped into two main categories, ``absolute'' (\ref{fig:fig_absolute_thermodynamic_cycle}) and ``relative'' \footnote{The distinction is a bit of a misnomer, since both compute ratios of partition functions relative to another state, and neither computes an absolute free energy.}(\ref{fig:fig_binding_thermodynamic_cycle}), which differ in whether they compute properties for a single molecule (absolute) or compare properties of different, usually closely related, molecules (relative).
To use binding as a concrete example, in absolute binding free energy calculations, one computes the binding free energy of a ligand to an individual receptor relative to a standard reference concentration.
In contrast, in relative binding free energy calculations, one compares the binding free energy of two related inhibitors to determine the potency difference.
\subsection*{Absolute and relative free energy calculations have some differences}
Many of the issues of setup and protocol for alchemical calculations are common, but some are different between absolute and relative calculations, so before treating the common elements we consider the protocol differences.

@@

\paragraph{Topologies} A critical first step in relative calculations is to select an approach to these calculations, determining whether to use a \emph{dual topology}, \emph{single topology}, or \emph{hybrid topology} approach to relative calculations.

The distinction between these can be illustrated by considering a hypothetical transformation from molecule A to molecule B, where both atoms share a common substructure but differ in their substituents; e.g. consider a transformation of benzene (A) to benzyl alcohol (B) fig.~\ref{fig:fig4_topology}.
In this case the common substructure is the benzene ring, though perhaps may be larger depending on how it is defined, as we discuss below.
In single topology calculations, the overall transformation is set up to involve as few additional atoms as possible, so benzene would be typically changed into benzyl alcohol by changing one of the hydrogens into a carbon. To this new carbon you also find non-interacting atoms called ``dummy atoms'' (retaining their bonded interactions but not interacting with the rest of the system), which will become the two hydrogens attached to the new carbon and the OH group. Bond parameters as well as partial charges are adjusted accordingly. 
Thus, in a single topology calculation, atoms may change their type so relatively few dummy atoms are created. This is illustrated in fig.~\ref{fig:fig4_topology} (a). 
In contrast, in a dual topology free energy calculation, no atoms are allowed to change type~\cite{shirts2012best}. This means that the benzene to benzyl alcohol transformation involves starting with benzene plus the non-interacting dummy atoms making up the hydroxy methyl group, then passing through an intermediate state where atoms which are becoming dummy atoms or ceasing being dummy atoms are partially interacting (this state may or may not be well defined~\cite{mobley2014blind}  and culminating in a state where benzyl alcohol is present along with the additional dummy atoms which was previously a corresponding hydrogen of the benzene. Fig.~\ref{fig:fig4_topology} (b) depicts how such a dual topology works. 
\begin{figure}[h!]
    \includegraphics[width=0.95\columnwidth]{figures/fig3_topol/Figure.pdf}
    \caption{top: single topology bottom: dual topology Figure adapted from \url{http://www.alchemistry.org/wiki/Constructing_a_Pathway_of_Intermediate_States}}
    \label{fig:fig_topology}
\end{figure} 
Hybrid topology calculations have seen much less use but essentially consist of two absolute free energy calculations in opposite directions at the same time (turning one molecule off while turning the other on)~\cite{jiang2019computing}.
At present, the most widely used approaches, such as in Schr\"{o}dinger's FEP+~\cite{wang2019protein} and in FESetup~\cite{loeffler2015fesetup} (for which calculations may be planned with Lead Optimization Mapper (LOMAP)~\cite{liu2013lead} seem to use single topology approaches, though some codes only support dual topology.
To our knowledge efficiency differences have not been thoroughly explored, though conventional wisdom suggests that fewer dummy atoms are better~\cite{liu2013lead,mobley2012perspective}.
